Carbonate chemistry at Ny-Ålesund

### 2022-08-22 awipev-CO2.Rmd
Updated air-sea CO2 flux

tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_remos at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_sf_4 ~ tmp$depth_group * tmp$montht))

@@

## Atmospheric CO2 Zeppelin
```{r atmospheric CO2 Zeppelin, echo=FALSE, message= FALSE, warning=FALSE}
# downloaded 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999
# value:units : ppm
# value:long_name : measured_mole_fraction_of_trace_gas_in_dry_air
atm_co2_old <-
  read_delim(file = paste0(path, "atmospheric_co2_zeppelin_old/co2_zep_surface-insitu_54_9999-9999_hourly.txt"),
  delim = " ",
  comment = "#",
  #skip = 182,
  col_names = TRUE,
  na = "-999.999",
  show_col_types = FALSE
  ) %>%
  dplyr::mutate(co2_air = value,
  datetime = as_datetime(paste(paste(year, month, day, sep="-"), paste(hour, minute, second, sep="-"), sep = " "))) %>%
  dplyr::select(datetime, co2_air)

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
pco2_fb_remos <- d_all %>% 
  dplyr::select(depth, monthd, pco2_fb_ctd_temp, pco2_remos_ctd_temp) %>% 
  dplyr::filter(depth < 2) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_ctd_temp, y = pco2_remos_ctd_temp, label = monthd)) +
  geom_point(col="blue", size = 0.5) +
  #geom_text() +
  xlim(180, 420) + ylim(180, 420) +
  labs(title = "pCO2 at ctd temperature",
       x=latex2exp::TeX("$pCO_{2}$ 11 m $\\phantom{x}(\\muatm)$"),
       y= latex2exp::TeX("$pCO_{2}$ REMOS at 0-2 m\\phantom{x}($\\muatm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 9) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
pco2_fb_remos
ggsave(filename = "figures/pco2_fb_remos.png", plot = pco2_fb_remos, width = 25, height = 25, units = "cm")
```

@@

##TO DO: check calculations by Miller et al. (2020); use formulation of Wanninkhof 2014?
## TO DO: Dobashi R. & Ho D. T., 2022. Air-sea gas exchange in a seagrass ecosystem. EGUsphere 
d_all <- d_all %>%
  dplyr::mutate(delta_pco2 = pco2_fb - co2_air, # ppm
                k600 = 0.266 * ws^2, #cm/hr Ho et al. (2006)
                Sc = 2073.1 - (125.62*temp_insitu_11m) + (3.6276*temp_insitu_11m^2) -
                  (0.043219*temp_insitu_ctd^3), #dimensionless
                k = k600 * (Sc/600)^-0.5, #cm/h
                alpha = seacarb::K0(S=sal_fb_mix,T=temp_insitu_11m,P=0,Patm=p_air/1013.25), #mol/(kg/atm)
                alpha = alpha * (seacarb::rho(S = sal_fb_mix, T = temp_insitu_11m, P = 0)/1000), #mol/(l/atm)
                f_air_sea = k * alpha * delta_pco2 # mmol/(m2 h)
                )

@@

air_sea_month <- d_all %>%
  dplyr::select(monthd, f_air_sea) %>%
  dplyr::group_by(monthd) %>%
  dplyr::summarise(f_air_sea = 24 * mean(f_air_sea, na.rm = TRUE)/1000) %>% #mol/(m2 d)
  ungroup() %>%
  dplyr::mutate(f_air_sea = days_in_month(monthd) * f_air_sea) %>% 
  dplyr::mutate(montht = month(monthd, label = TRUE, abbr = TRUE))
#air_sea_month
p_f_air_sea_m <- ggplot(data = air_sea_month) +
  geom_point(aes(x = montht, y = f_air_sea), col = "blue", size = 2, na.rm = TRUE) +
  #scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  #lims(y = c(-30, 0)) +
  labs(title = "Air-sea CO2 flux (per month)", x = "Month", y = "CO2 (mmol/(m2 month))") +
  Mytheme(size_labs = 8)
p_f_air_sea_m
ggsave(filename = "figures/f_air_sea_m.png", plot = p_f_air_sea_m, width = 20, units = "cm")

p <- p_f_air_sea / p_f_air_sea_d / p_f_air_sea_m
ggsave(filename = "figures/air-sea_CO2_flux.png", plot = p, width = 20, units = "cm")
knitr::include_graphics("figures/air-sea_CO2_flux.png")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

The annual air-sea CO2 flux is
`r round(sum(air_sea_month$f_air_sea), digits = 0)` mol/(m2 year).

>>>>

tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_remos at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_4 ~ tmp$depth_group * tmp$montht))

@@

## Atmospheric CO2 Zeppelin
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# downloaded 2020-08-19 from https://gaw.kishou.go.jp/search/file/0054-6001-1001-01-01-9999
# value:units : ppm
# value:long_name : measured_mole_fraction_of_trace_gas_in_dry_air
atm_co2_old <-
  read_delim(file = paste0(path, "atmospheric_co2_zeppelin_old/co2_zep_surface-insitu_54_9999-9999_hourly.txt"),
  delim = " ",
  comment = "#",
  #skip = 182,
  col_names = TRUE,
  na = "-999.999",
  show_col_types = FALSE
  ) %>%
  dplyr::mutate(co2_air = value,
  datetime = as_datetime(paste(paste(year, month, day, sep="-"), paste(hour, minute, second, sep="-"), sep = " "))) %>%
  dplyr::select(datetime, co2_air)

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
p_pco2_fb_remos <- d_all %>% 
  dplyr::select(depth, monthd, pco2_fb_ctd_temp, pco2_remos_ctd_temp) %>% 
  dplyr::filter(depth < 4) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_ctd_temp, y = pco2_remos_ctd_temp, label = monthd)) +
  geom_point(col="blue", size = 0.5) +
  #geom_text() +
  xlim(250, 400) + ylim(270, 360) +
  labs(title = "pCO2 at ctd temperature",
       x=latex2exp::TeX("$pCO_{2}$ 11 m $\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("$pCO_{2}$ REMOS at 0-4 m\\phantom{x}($\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(mapping = use_label(c("R2", "eq"))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 9) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
p_pco2_fb_remos
ggsave(filename = "figures/pco2_fb_remos.png", plot = p_pco2_fb_remos, width = 25, height = 25, units = "cm")

# Same thing but monthly median; code different because one wants to calculate delta
tmp <- d_all %>% 
  dplyr::select(depth, montht, pco2_fb_ctd_temp, pco2_remos_ctd_temp) %>% 
  dplyr::filter(depth < 4) %>% 
  drop_na() %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(
    pco2_fb_ctd_temp = median(pco2_fb_ctd_temp, na.rm = TRUE),
    pco2_remos_ctd_temp = median(pco2_remos_ctd_temp, na.rm = TRUE), 
    delta_pco2_fb_remos_monthly = pco2_remos_ctd_temp - pco2_fb_ctd_temp
  )
p_pco2_fb_remos_month <- tmp %>% 
  ggplot(aes(x = pco2_fb_ctd_temp, y = pco2_remos_ctd_temp, label = montht)) +
  geom_point(col="blue", size = 0.5) +
  #geom_text() +
  xlim(250, 400) + ylim(250, 400) +
  labs(title = "pCO2 at ctd temperature",
       x=latex2exp::TeX("$pCO_{2}$ 11 m $\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("$pCO_{2}$ REMOS at 0-4 m\\phantom{x}($\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(mapping = use_label(c("R2", "eq"))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 9) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
p_pco2_fb_remos_month
ggsave(filename = "figures/pco2_fb_remos_monthly.png", plot = p_pco2_fb_remos_month, width = 25, height = 25, units = "cm")

# combine
g <- p_pco2_fb_remos / p_pco2_fb_remos_month
g <- g + patchwork::plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/pco2_compare.png", plot = g)
g
```
The relationship between pCO2 between 0 and 4 m on the REMOS (estimated from pH and S-derived AT) and pCO2 measured at 11 m (both pCO2 at CTD temperature) is much nicer using monthly median values (b) than using all data (a). The mean and median difference between the two pCO2s are `r round(summary(tmp$delta_pco2_fb_remos_monthly), digits=0)`

@@

##TO DO: check calculations by Miller et al. (2020); use formulation of Wanninkhof 2014?
## TO DO: Dobashi R. & Ho D. T., 2022. Air-sea gas exchange in a seagrass ecosystem. EGUsphere 
d_all <- d_all %>%
  dplyr::mutate(delta_pco2 = pco2_fb_ctd_temp - co2_air, # ppm
                k600 = 0.266 * ws^2, #cm/hr Ho et al. (2006)
                Sc = 2073.1 - (125.62*temp_insitu_ctd) + (3.6276*temp_insitu_ctd^2) -
                  (0.043219*temp_insitu_ctd^3), #dimensionless
                k = k600 * (Sc/600)^-0.5, #cm/h
                alpha = seacarb::K0(S=sal_insitu_ctd,T=temp_insitu_ctd,P=0,Patm=p_air/1013.25), #mol/(kg/atm)
                alpha = alpha * (seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = 0)/1000), #mol/(l/atm)
                f_air_sea = k * alpha * delta_pco2 # mmol/(m2 h)
                )

@@

air_sea_month <- d_all %>%
  dplyr::select(monthd, f_air_sea) %>%
  dplyr::group_by(monthd) %>%
  dplyr::summarise(f_air_sea = 24 * mean(f_air_sea, na.rm = TRUE)/1000) %>% #mol/(m2 d)
  ungroup() %>%
  dplyr::mutate(f_air_sea = days_in_month(monthd) * f_air_sea) %>% 
  dplyr::mutate(montht = month(monthd, label = TRUE, abbr = TRUE))
#air_sea_month
p_f_air_sea_month <- ggplot(data = air_sea_month) +
  geom_point(aes(x = montht, y = f_air_sea), col = "blue", size = 2, na.rm = TRUE) +
  #scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  lims(y = c(-30, 0)) +
  labs(title = "Air-sea CO2 flux (per month)", x = "Month", y = "CO2 (mmol/(m2 month))") +
  Mytheme(size_labs = 8)
p_f_air_sea_month
ggsave(filename = "figures/f_air_sea_m.png", plot = p_f_air_sea_month, width = 20, units = "cm")

p <- p_f_air_sea / p_f_air_sea_d / p_f_air_sea_month
ggsave(filename = "figures/air-sea_CO2_flux.png", plot = p, width = 20, units = "cm")
knitr::include_graphics("figures/air-sea_CO2_flux.png")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

```{r air-sea CO2 with corrected data, echo=FALSE, message= FALSE, warning=FALSE}
# Same as above but we correct pCO2 by adding 17 uatm as we show above that pCO2 at 11 m
# underestimates pCO2 at 0-4m (hence overestimates CO2 uptake)
# round(summary(tmp$delta_pco2_fb_remos_monthly))[[4]]
d_all <- d_all %>%
  dplyr::mutate(delta_pco2_corr = (pco2_fb_ctd_temp + 17) - co2_air, # ppm
                k600 = 0.266 * ws^2, #cm/hr Ho et al. (2006)
                Sc = 2073.1 - (125.62*temp_insitu_ctd) + (3.6276*temp_insitu_ctd^2) -
                  (0.043219*temp_insitu_ctd^3), #dimensionless
                k = k600 * (Sc/600)^-0.5, #cm/h
                alpha = seacarb::K0(S=sal_insitu_ctd,T=temp_insitu_ctd,P=0,Patm=p_air/1013.25), #mol/(kg/atm)
                alpha = alpha * (seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = 0)/1000), #mol/(l/atm)
                f_air_sea_corr = k * alpha * delta_pco2_corr # mmol/(m2 h)
                )

# p_f_air_sea <- ggplot(data = d_all) +
#   geom_point(aes(x = datetime, y = f_air_sea), col = "blue", size = 0.15, na.rm = TRUE) +
#   labs(title = "Air-sea CO2 flux (all data)", x = "Time", y = "CO2 (mmol/(m2 h))") +
#   Mytheme(size_labs = 8)
# p_f_air_sea
# ggsave(filename = "figures/p_f_air_sea.png", plot = p_f_air_sea, width = 20, units = "cm")

# air_sea_day_d <- d_all %>%
#   dplyr::select(datetime, f_air_sea_corr) %>%
#   dplyr::group_by(lubridate::yday(datetime)) %>%
#   dplyr::summarise(f_air_sea_corr = 24 * mean(f_air_sea_corr, na.rm = TRUE)/1000) %>% #mol/(m2 d)
#   ungroup() %>%
#   dplyr::rename(day = 1)

# p_f_air_sea_d <- ggplot(data = air_sea_day) +
#   geom_point(aes(x = day, y = f_air_sea), col = "blue", size = 0.15, na.rm = TRUE) +
#   labs(title = "Air-sea CO2 flux (per day)", x = "Julian day", y = "CO2 (mmol/(m2 d))") +
#   Mytheme(size_labs = 8)
# p_f_air_sea_d
# ggsave(filename = "figures/f_air_sea_d.png", plot = p_f_air_sea_d, width = 20, units = "cm")
# 
air_sea_corr_month <- d_all %>%
  dplyr::select(monthd, f_air_sea_corr) %>%
  dplyr::group_by(monthd) %>%
  dplyr::summarise(f_air_sea_corr = 24 * mean(f_air_sea_corr, na.rm = TRUE)/1000) %>% #mol/(m2 d)
  ungroup() %>%
  dplyr::mutate(f_air_sea_corr = days_in_month(monthd) * f_air_sea_corr) %>% 
  dplyr::mutate(montht = month(monthd, label = TRUE, abbr = TRUE))
#air_sea_month
p_f_air_sea_corr_month <- ggplot(data = air_sea_corr_month) +
  geom_point(aes(x = montht, y = f_air_sea_corr), col = "blue", size = 2, na.rm = TRUE) +
  #scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  lims(y = c(-30, 0)) +
  labs(title = "Air-sea CO2 flux (per month)", x = "Month", y = "CO2 (mmol/(m2 month))") +
  Mytheme(size_labs = 8)
p_f_air_sea_corr_month
ggsave(filename = "figures/f_air_sea_corr_month.png", plot = p_f_air_sea_corr_month, width = 20, units = "cm")
# combine without and with correction
p <- p_f_air_sea_month / p_f_air_sea_corr_month
ggsave(filename = "figures/air-sea_CO2_flux_comp.png", plot = p, width = 20, units = "cm")
knitr::include_graphics("figures/air-sea_CO2_flux_comp.png")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

The annual air-sea CO2 flux is
`r round(sum(air_sea_month$f_air_sea), digits = 0)` and `r round(sum(air_sea_corr_month$f_air_sea_corr), digits = 0)` mol/(m2 year) without and with correction, respectively.
